Toronto Paramedic Services

### 2024-09-26 paper/paper.qmd
Got a working map done - just need to change labels on legend.

ggplot(data, aes(x = Year, y = Total_Emergency_Calls)) + 
  geom_line(linewidth = 0.5, color="black") + geom_point(colour = "black") +
  ylim(200000,300000) + theme_light() + 
  ggtitle("EMS Demand From 2017 to 2022") +
  labs(x = "Year", y = "Demand for EMS", caption = "Data Source: Toronto Paramedic Services") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```
@fig-totalemsdemand displays the total number of emergency calls for each year. The total number of emergency calls appears to increase in 2020 after a notable decrease in 2019. This pattern likely reflects the the onset of the covid-19 pandemic. Though the number of emergency calls appears to fluctuate over the years, the change is not drastic, indicating that there has consistently been a heightened demand for EMS within Toronto.

@@

data |>
  pivot_longer(
    cols = c("Total_Emergency_Calls", "Medical_Emergency_Calls", "Vehicle_Accident_Calls", "Emergency_Transfer_Calls", "Fire_Calls", "Airport_Standby_Calls"),
    names_to = "series",
    values_to = "value"
  ) |>
  ggplot(mapping = aes(x = Year, y = value, color = series)) +
  geom_line() +
  geom_point() +
  theme_light() +
  labs(
    x = "Year", color = "Incident Type",
    caption = "Data Source: Toronto Paramedic Services"
  ) +
  scale_color_brewer(palette = "Set3", labels = c("Airport Standbys", "Emergency Transfers", "Fires", "Medical Emergencies", "Total Emergency Calls", "Vehicle Accidents")) +
  ggtitle("EMS Demand (2017-2022) Based on Incident Type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylim(0,300000) + scale_y_continuous(name = "Demand for EMS", labels = scales::comma)
```
```{r}
#| echo: false 
#| message: false
#| label: fig-incidentemsdemandzoom
#| fig-cap: EMS Demand for Each Year between 2017 and 2022

@@

data |>
  pivot_longer(
    cols = c("Vehicle_Accident_Calls", "Emergency_Transfer_Calls", "Fire_Calls", "Airport_Standby_Calls"),
    names_to = "series",
    values_to = "value"
  ) |>
  ggplot(mapping = aes(x = Year, y = value, color = series)) +
  geom_line() +
  geom_point() +
  theme_light() +
  labs(
    x = "Year", color = "Incident Type",
    caption = "Data Source: Toronto Paramedic Services"
  ) +
  scale_color_brewer(palette = "Set3", labels = c("Airport Standbys", "Emergency Transfers", "Fires", "Vehicle Accidents")) +
  ggtitle("EMS Demand (2017-2022) Based on Incident Type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylim(0,20000) + scale_y_continuous(name = "Demand for EMS", labels = scales::comma)
```

>>>>

ggplot(data, aes(x = Year, y = Total_Emergency_Calls)) + 
  geom_line(linewidth = 0.5, colour="black") + geom_point(colour = "black") +
  ylim(200000,300000) + theme_light() + 
  ggtitle("EMS Demand From 2017 to 2022") +
  labs(x = "Year", y = "Demand for EMS", caption = "Data Source: Toronto Paramedic Services") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```
@fig-totalemsdemand displays the total number of emergency calls for each year. The total number of emergency calls appears to increase in 2020 after a notable decrease in 2019. This pattern likely reflects the the onset of the covid-19 pandemic. Though the number of emergency calls appears to fluctuate over the years, the change is not drastic, indicating that there has consistently been a heightened demand for EMS within Toronto.

@@

data |>
  pivot_longer(
    cols = c("Total_Emergency_Calls", "Medical_Emergency_Calls", "Vehicle_Accident_Calls", "Emergency_Transfer_Calls", "Fire_Calls", "Airport_Standby_Calls"),
    names_to = "series",
    values_to = "value"
  ) |>
  ggplot(mapping = aes(x = Year, y = value, colour = series)) +
  geom_line() +
  geom_point() +
  theme_light() +
  labs(
    x = "Year", colour = "Incident Type",
    caption = "Data Source: Toronto Paramedic Services"
  ) +
  scale_colour_brewer(palette = "Set3", labels = c("Airport Standbys", "Emergency Transfers", "Fires", "Medical Emergencies", "Total Emergency Calls", "Vehicle Accidents")) +
  ggtitle("EMS Demand (2017-2022) Based on Incident Type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylim(0,300000) + scale_y_continuous(name = "Demand for EMS", labels = scales::comma)
```
```{r}
#| echo: false 
#| message: false
#| label: fig-incidentemsdemandzoom
#| fig-cap: EMS Demand for Each Year between 2017 and 2022

@@

data |>
  pivot_longer(
    cols = c("Vehicle_Accident_Calls", "Emergency_Transfer_Calls", "Fire_Calls", "Airport_Standby_Calls"),
    names_to = "series",
    values_to = "value"
  ) |>
  ggplot(mapping = aes(x = Year, y = value, colour = series)) +
  geom_line() +
  geom_point() +
  theme_light() +
  labs(
    x = "Year", colour = "Incident Type",
    caption = "Data Source: Toronto Paramedic Services"
  ) +
  scale_colour_brewer(palette = "Set3", labels = c("Airport Standbys", "Emergency Transfers", "Fires", "Vehicle Accidents")) +
  ggtitle("EMS Demand (2017-2022) Based on Incident Type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylim(0,20000) + scale_y_continuous(name = "Demand for EMS", labels = scales::comma)
```

## Mapping EMS Demand

```{r}
#| echo: false 
#| message: false
#| label: fig-map
#| fig-cap: EMS Demand Across Toronto Forward Sortation Areas (Based on data from 2022)

library(maps)
library(sf)

# Reading in the shapefile

toronto_forward_sortation_areas <- st_read("~/Toronto_Paramedic_Services/data/TorontoForwardSortationAreas")
#view(toronto_forward_sortation_areas)

# Reading in the paramedic services data

map_data <- read_csv(here::here("data/analysis_data/analysis_data_FOR_MAP.csv"))
#view(map_data)

# Test to see if shapefile produces a map

#ggplot(toronto_forward_sortation_areas) + 
  #geom_sf()

# Merge map_data and toronto_forward_sortation_areas by the column "CFSAUID"

merged_map_data <- merge(toronto_forward_sortation_areas, map_data, by = "CFSAUID")
#view(merged_map_data)

# Map the merged data

ggplot(merged_map_data) + 
  geom_sf(aes(fill = "EMS Demand Driver")) +
  labs(title = "EMS Demand Drivers By Forward Sortation Area (FSA)",
         x = "Longitude", y="Latitude", 
       caption = "Data Source: Open Canada 2021 Census - Boundary Files & Toronto Paramedic Services",
       labels = c("Medical Emergencies")) +
  theme_light() + theme(plot.title = element_text(hjust = 0.5))
